The Analog Method for Precipitation Downscaling: Finding Better Analog Situations at a Sub-Daily Time Step

### 2016-11-15 Moving temporal window - HESS/template.tex
HESS paper under corrections.

\begin{abstract}
The Analog Method (AM) aims at predicting local weather variables (predictands), such as precipitations, by means of a statistical relationship with predictors at a synoptic scale. The analogy is generally assessed in the first place on the geopotential field by mean of a comparison of the gradients, in order to sample the days with a similar atmospheric circulation.

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day, for both the target day and the candidate analogs. The constraint being the use of daily time series, due to the length of available archives they provide, and the unavailability of equivalent archives at a finer time step. However, it is unlikely that the best analogy happens at the very same hour, but it may occur at a different time of the day. In order to assess the potential of finding better analogs at a different hour, a moving time window (MTW) has been introduced on a reduced archive of hourly precipitation totals.

@@

The Analog Method (AM) is based on the principle that two relatively similar synoptic situations may produce relatively similar local effects \citep{Lorenz1956, Lorenz1969}. Multiple variations of the methods exist and listings can be found in \citet{BenDaoud2016}. The versions that will be considered here are often used as references or benchmarks for various improvements.

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent at a finer resolution. Therefore, the analog situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day, otherwise one would not know what precipitation values to assign to them. However, it can be expected that the analogy of the synoptic situations does not occur systematically at the same time of the day, and that better candidates can be found by shifting to a different hour. On this assumption, a moving time window (MTW) was introduced to allow searching for candidates at different hours of the day.

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland \cite[see also][]{Horton2012a}. Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The time series (on the period 1982--2007) come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Fig. \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The results will thereafter be presented arbitrarily for the Ulrichen station, but they are equivalent for all stations. The precipitation time series were then aggregated by means of a moving 24~h-total (see next section).

@@

The predictand, here a 24~h precipitation total, results from meteorological situations that are continuous transitions between changing circulation patterns. Thus, the observed predictors fields, extracted from reanalysis datasets, are considered at different time of the day and at different heights. In our case, the geopotential height is considered at 1000~hPa (Z1000) at 12~h~UTC and 500~hPa (Z500) at 24~h~UTC. The selection of the observation time of the predictor was found by \citet{Bontron2004} to have a significant influence.

@@

To calibrate the method, the commonly used technique is a semi-automatic sequential procedure elaborated by \citet{Bontron2004}.

The target situations and their corresponding observed precipitation values (used for validation) do not change, because the prediction is always established for a fixed period of the target day (6--30~h), as before. The difference is that the candidates are 4~times more numerous (Fig. \ref{fig:principle}).

Based on some preliminary tests (not shown), no constraint was added in order to restrict the selection of multiple analogs within the same candidate dates.

@@

\begin{equation}
\label{eq:CRPSS}
S_{\text{SCRP}} = \frac{S_{\text{CRP}}-S_{\text{CRP}}^{r}}{S_{\text{CRP}}^{p}-S_{\text{CRP}}^{r}} = 1-\frac{S_{\text{CRP}}}{S_{\text{CRP}}^{r}}
\end{equation}
where $S_{\text{CRP}}^{r}$ is the $S_{\text{CRP}}$ value for the reference (climatological distribution) and $S_{\text{CRP}}^{p}$ would be the one for a perfect prediction (which implies $S_{\text{CRP}}^{p}~=~0$). A better prediction is characterized by an increase in CRPSS: $S_{\text{SCRP}}~=~1$ for a perfect prediction, and $S_{\text{SCRP}}~<~0$ for a prediction with a lower skill than the reference.

@@

@@

Figure \ref{fig:changes_S1_analogs} presents the changes in the distributions of the S1 criterion for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analog situations for the Ulrichen station on the whole calibration period (1961-2008), due to introduction of the MTW. The precipitation target remains as before, that is centered on 18~h~UTC (6~h~UTC to 6~h~UTC the next day). The shapes of the distributions of the conventional approach and the MTW are similar, but the values of the analogy criteria are now reduced (shifted to the left), and therefore better. For analogs with higher ranks (e.g. $20^{th}$ or $40^{th}$), the difference between the two distributions is larger than for the first ranks, which means that the improvement increases with the rank of the analogs. This is due to the accumulation of improvements brought by the new analog situations in the selection.

The improvements of the S1 criteria are summarized in Fig. \ref{fig:changes_S1}, which shows (top) quantiles of the S1 criteria according to the analog rank for the conventional method and the MTW, and (bottom) quantiles of the relative reduction. This confirms that all quantile seem similarly reduced (S1 distributions keep their shape), and that this improvement is constantly increasing from the first to the last analog (Fig. \ref{fig:changes_S1} bottom).

@@

The S1 criteria was previously found to be improved to a greater extent for the most dynamic situations related with higher precipitation values (Sect. \ref{sec:influence_precip}). The changes in terms of performance scores will now be assessed regarding precipitation thresholds. Fig. \ref{fig:changes_CRPS_precip_threshold} synthesizes these differences for the Ulrichen station, other stations having the same behavior.

@@

On the contrary, the summer months have a preference for the initial temporal window (Z500 24~h \& Z1000 12~h), due to more pronounced diurnal effects which reduces the potential for improvement of the criteria. Other seasons are between these two extremes, which is consistent with their respective improvements. This seasonal effect was observed for each station in a very similar way, and generally even with a slightly larger amplitude than for Ulrichen (see example of the Visp station in Fig. \ref{fig:hours_selection_per_season_Visp}).

@@

\subsection{Is a simple reconstructed archive useful for the MTW ?}

@@

\subsection{Can a proxy variable help reconstruct a better archive ?}

@@

The importance of the quality of the precipitation archive was also demonstrated, as too simplistic reconstructions of 6-hourly time series lead to a loss of all the improvement brought by better analog situations identified by the MTW. The precipitation prediction is improved only when the precipitation chronology is close to the accurate one. Attempts to reconstruct longer time series based on a simplistic proportional distribution or using meteorological variables from the NCEP reanalysis 1 dataset as proxy did not succeed. Other reanalysis datasets could eventually preform better.

Perspectives of using the MTW rely on the availability of long precipitation series at a sub-daily time step and with high accuracy. The first source of such data is the simple fact that with time, these archives of high temporal resolution precipitation data are getting longer. The second source is the establishment of precipitation reanalysis at a country scale, or the use of reanalysis-driven RCMs (Regional Climate Models) or LAMs (Limited Area Models) on a long period. When such precipitation data will be available, the use of a MTW in the AM should be considered for any application, especially when the prediction quality of extreme events is important.

@@

\begin{figure}[htb]
	\includegraphics[width=8.3cm]{figures/hours_selection_per_season_Visp.pdf}
	\caption{Distribution of the predictors hours in the selected analog dates depending on the season, for the Visp station.}
	\label{fig:hours_selection_per_season_Visp}
\end{figure}

>>>>

\begin{abstract}
The Analog Method (AM) aims at predicting local weather variables (predictands), such as precipitation, by means of a statistical relationship with predictors at a synoptic scale. The analogy is generally first assessed on the geopotential field gradients, in order to sample days with a similar atmospheric circulation. Other predictors, such as moisture variables, can also be added in a successive level of analogy.

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day, for both the target day and the candidate analogs. The constraint is the use of daily time series, due to the length of available archives they provide, and the unavailability of equivalent archives at a finer time step. However, it is unlikely that the best analogy happens at the very same hour, but it may occur at a different time of the day. In order to assess the potential of finding better analogs at a different hour, a moving time window (MTW) has been introduced on a reduced archive of hourly precipitation totals.

@@

The Analog Method (AM) is based on the principle that two relatively similar synoptic situations may produce relatively similar local effects \citep{Lorenz1956, Lorenz1969}. Multiple variations of the methods exist and listings can be found in \citet{BenDaoud2016}.

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent at a finer resolution. Therefore, the analog situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day. However, it can be expected that the analogy of the synoptic situations does not occur systematically at the same time of the day, and that better candidates can be found by shifting to a different hour. On this assumption, a moving time window (MTW) was introduced to allow searching for candidates at different hours of the day.

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland \cite[see also][]{Horton2012a}. Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The time series (on the period 1982--2007) come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Fig. \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The results will thereafter be presented arbitrarily for the Ulrichen station, but they are equivalent for all stations.

@@

The predictand, here a 24~h precipitation total (6~h~UTC to 6~h~UTC the next day), results from meteorological situations that are continuous transitions between changing circulation patterns. Thus, the observed predictors fields, extracted from reanalysis datasets, are considered at different hours and at different heights. In our case, the geopotential height is considered at 1000~hPa (Z1000) at 12~h~UTC and 500~hPa (Z500) at 24~h~UTC. The selection of the time of the day at which the predictors are selected was found by \citet{Bontron2004} to have a significant influence.

@@

@@

\begin{equation}
\label{eq:CRPSS}
S_{\text{SCRP}} = \frac{S_{\text{CRP}}-S_{\text{CRP}}^{r}}{S_{\text{CRP}}^{p}-S_{\text{CRP}}^{r}} = 1-\frac{S_{\text{CRP}}}{S_{\text{CRP}}^{r}}
\end{equation}
where $S_{\text{CRP}}^{r}$ is the $S_{\text{CRP}}$ value for the reference (climatological distribution \textendash\ i.e. the distribution of all precipitation values from the corresponding archive) and $S_{\text{CRP}}^{p}$ would be the one for a perfect prediction (which implies $S_{\text{CRP}}^{p}~=~0$). A better prediction is characterized by an increase in CRPSS: $S_{\text{SCRP}}~=~1$ for a perfect prediction, and $S_{\text{SCRP}}~<~0$ for a prediction with a lower skill than the reference.

@@

The target situations and their corresponding observed precipitation values do not change, because the prediction is always established for a fixed period of the target day (6~h~UTC to 6~h~UTC the next day), as before. The difference is that the candidates are 4~times more numerous (Fig. \ref{fig:principle}).

No constraint was added in order to restrict the selection of multiple analogs within the same candidate dates.

To calibrate the method, the semi-automatic sequential procedure elaborated by \citet{Bontron2004} was used \cite[see also ][]{BenDaoud2016}. The analogy levels (eg. the atmospheric circulation or moisture index) are calibrated sequentially. The parameters calibrated by this approach, for every level of analogy, are the spatial windows on which the predictors are compared and the number of analogs. The procedure consists of the following steps \citep{Bontron2004}:

\begin{enumerate}
	\item Manual selection of the following parameters:
	\begin{enumerate}
		\item meteorological variable (e.g. geopotential height, temperature, relative humidity, etc.),
		\item pressure level (e.g. 500 hPa, ...),
		\item temporal window (hour of the day \textendash\ e.g. 6, or 12~h~UTC),
		\item initial analog numbers (e.g. $N_{1}=50$).
	\end{enumerate}

	\item For every level of analogy:
	\begin{enumerate}
		\item Identification of the most skilled unitary cell (1~point for moisture variables and 4 for the geopotential height when using the S1 criteria) over a large domain. Every point (or cell) of the full domain is assessed jointly on all predictors of the level of analogy (consisting generally of the same variable, but on different pressure levels and at different hours).
		\item From this most skilled point, the spatial window is expanded by successive iterations in the direction of greater performance gain. The spatial window grows up by repeating the previous steps, until no improvement is reached.
		\item The number of analog situations is then reconsidered and optimized for the current level of analogy.
		\item A new level of analogy can then be added, based on other variables (such as the moisture index) at chosen pressure levels and hours of the day. The number of analogs for the next level of analogy is initiated at a chosen value. Then, the procedure starts again from step (a) for the new level. The parameters calibrated on the previous analog levels are fixed and do not change (except the numbers of analogs of every level of analogy, that are optimized once again at the final stage). 
	\end{enumerate}
	\item Finally, the numbers of analogs are re-assessed for the different levels of analogy. This is done iteratively by varying the number of analogs of each level in a systematic way.
\end{enumerate}

The calibration is thus done in successive steps, on a limited number of parameters. Previously calibrated parameters are generally not reassessed.

@@

Figure \ref{fig:changes_S1_analogs} presents the changes in the distributions of the S1 criterion for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analog situations for the Ulrichen station on the whole calibration period (1961-2008), due to introduction of the MTW. The precipitation target remains as before, that is centered on 18~h~UTC (6~h~UTC to 6~h~UTC the next day). The shapes of the distributions of the conventional approach and the MTW are similar, but the values of the analogy criteria are now reduced (shifted to the left), and therefore better. For analogs with higher ranks (e.g. $20^{th}$ or $40^{th}$), the difference between the two distributions is larger than for the first ranks, which means that the improvement increases with the rank of the analogs.

The improvements of the S1 criteria are summarized in Fig. \ref{fig:changes_S1}, which shows (top) quantiles of the S1 criteria according to the analog rank for the conventional method and the MTW, and (bottom) quantiles of the relative reduction. This confirms that all quantiles are similarly reduced (S1 distributions keep their shape), and that this improvement is constantly increasing from the first to the last analog (Fig. \ref{fig:changes_S1} bottom).

@@

The S1 criteria was previously found to be improved to a greater extent for higher precipitation values (Sect. \ref{sec:influence_precip}). The changes in terms of performance scores will now be assessed regarding precipitation thresholds. Fig. \ref{fig:changes_CRPS_precip_threshold} synthesizes these differences for the Ulrichen station, other stations having the same behavior.

@@

On the contrary, the summer months have a preference for the initial temporal window (Z500 24~h \& Z1000 12~h), due to more pronounced diurnal effects which reduces the potential for improvement of the criteria. Other seasons are between these two extremes, which is consistent with their respective improvements. This seasonal effect was observed for each station in a very similar way, and generally even with a slightly larger amplitude than for Ulrichen.

@@

\subsection{Is a simple reconstructed precipitation archive useful for the MTW ?}

@@

\subsection{Can a proxy variable help reconstruct a better precipitation archive ?}

@@

The importance of the quality of the precipitation archive was also demonstrated, as too simplistic reconstructions of 6-hourly time series lead to a loss of all the improvement brought by better analog situations identified by the MTW. The precipitation prediction is improved only when the precipitation chronology is close to the accurate one. Attempts to reconstruct longer time series based on a simplistic proportional distribution or using meteorological variables from the NCEP reanalysis 1 dataset as proxy did not succeed. Other reanalysis datasets could eventually perform better.

Perspectives of using the MTW rely on the availability of long precipitation series at a sub-daily time step and with high accuracy. First, these archives of high temporal resolution precipitation data are increasing now with time. The second source is the establishment of precipitation reanalysis at a country scale, or the use of reanalysis-driven RCMs (Regional Climate Models) or LAMs (Limited Area Models) on a long period. When such precipitation data will be available, the use of a MTW in the AM should be considered for any application, especially when the prediction quality of extreme events is important.

@@
